---
title: "Case Study: How Does a Bike-Share Navigate Speedy Success?"
author: "Markos Tsinganis"
format: html
editor: visual
---

## Case Study Scenario

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company's future success depends on maximizing the number of annual memberships. Therefore, your team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, your team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data visualizations.

### Characters and teams

-   Cyclistic: A bike-share program that features more than 5,800 bicycles and 600 docking stations. Cyclistic sets itself apart by also offering reclining bikes, hand tricycles, and cargo bikes, making bike-share more inclusive to people with disabilities and riders who can't use a standard two-wheeled bike. The majority of riders opt for traditional bikes; about 8% of riders use the assistive options. Cyclistic users are more likely to ride for leisure, but about 30% use them to commute to work each day.
-   Lily Moreno: The director of marketing and your manager. Moreno is responsible for the development of campaigns and initiatives to promote the bike-share program. These may include email, social media, and other channels.
-   Cyclistic marketing analytics team: A team of data analysts who are responsible for collecting, analyzing, and reporting data that helps guide Cyclistic marketing strategy. You joined this team six months ago and have been busy learning about Cyclistic's mission and business goals --- as well as how you, as a junior data analyst, can help Cyclistic achieve them.
-   Cyclistic executive team: The notoriously detail-oriented executive team will decide whether to approve the recommended marketing program.

### About the company

In 2016, Cyclistic launched a successful bike-share offering. Since then, the program has grown to a fleet of 5,824 bicycles that are geotracked and locked into a network of 692 stations across Chicago. The bikes can be unlocked from one station and returned to any other station in the system anytime.

Until now, Cyclistic's marketing strategy relied on building general awareness and appealing to broad consumer segments. One approach that helped make these things possible was the flexibility of its pricing plans: single-ride passes, full-day passes, and annual memberships. Customers who purchase single-ride or full-day passes are referred to as casual riders. Customers who purchase annual memberships are Cyclistic members.

Cyclistic's finance analysts have concluded that annual members are much more profitable than casual riders. Although the pricing flexibility helps Cyclistic attract more customers, Moreno believes that maximizing the number of annual members will be key to future growth. Rather than creating a marketing campaign that targets all-new customers, Moreno believes there is a very good chance to convert casual riders into members. She notes that casual riders are already aware of the Cyclistic program and have chosen Cyclistic for their mobility needs.

Moreno has set a clear goal: Design marketing strategies aimed at converting casual riders into annual members. In order to do that, however, the marketing analyst team needs to better understand how annual members and casual riders differ, why casual riders would buy a membership, and how digital media could affect their marketing tactics. Moreno and her team are interested in analyzing the Cyclistic historical bike trip data to identify trends.

## 1. Ask

Financial analysts at Cyclistic have determined that annual members are more profitable than casual riders. The director of marketing believes future growth depends on maximizing the number of annual members. Our business problem is how to convince more potential clients to purchase an annual membership, rather than use our services for casual rides. In order to solve this business problem we need to design a new marketing strategy that will be informed by trends in the historical bike trip data. The current report has the following business task:

**How do annual members and casual riders use Cyclistic bikes differently?**

### Key stakeholders

## 2. Prepare

I will be using [Cyclistic's historical trip data](https://divvy-tripdata.s3.amazonaws.com/index.html) that have been made available by Motivate International Inc. under this [license](https://ride.divvybikes.com/data-license-agreement). The license includes the right to "access, reproduce, analyze, copy, modify, distribute in your product or service and use the Data for any lawful purpose". The bike trip data is organized in a directory containing zipped .csv files. Based on the file names we can deduce that they cover the period 2013 to 2022, with data from the last two years being published on a monthly basis. I will limit myself to the data from the most recent 12 months available (November 2021 - October 2022). The files in this time range use a consistent naming format ('YYYYMM-divvy-tripdata.zip') which makes it easy to download the files programmatically. As a result, I need to download 12 .zip files containing 12 separate .csv files.

A brief look at the file sizes suggests that importing these into RStudio and combining them in one dataframe can potentially crash the program or slow it down considerably. I therefore choose to avoid importing all the data in RStudio, but instead download the .csv files in my computer and save them as separate tables in a new SQLite database. I then use R to send SQL queries to the database to retrieve only the data I need for a particular analysis.

-   Where is your data located?
-   How is the data organized?
-   Are there issues with bias or credibility in this data? Does your data ROCCC?
-   How are you addressing licensing, privacy, security, and accessibility?
-   How did you verify the data's integrity?
-   How does it help you answer your question?
-   Are there any problems with the data?

Load required packages

```{r load libraries}
library(dplyr)
library(here)
library(stringr)
library(tidyr)
library(readr)
```


```{r}
csv_sizes <- tibble(file = list.files(here("raw_data")),
                    size = file.size(list.files(here("raw_data"),
                                                full.names = TRUE))) %>% 
               mutate(size_mb = size / 1048576)

library(ggplot2)
library(stringr)

my_labels <- sapply(csv_sizes$file, function(m) str_replace(m,
                                                pattern = "(.{4})(.{2})(.*)",
                                                replacement = paste("\\1-\\2")))

csv_sizes %>% 
    ggplot(aes(x = file, y = size_mb, group = 1)) + geom_line() +
    labs(title = paste0("Total size of files is ",
                        csv_sizes %>% summarize(round(sum(size_mb), 0)) %>% 
                            pull(), " MB")) +
    theme(axis.text.x = element_text(angle = 45)) +
    scale_x_discrete(labels = my_labels)


```



## 3. Process

## 4. Analyze

**knitr::opts_chunk\$set(fig.path = here::here("figs"))**

## 5. Share

## 6. Act
